Phenology and Mycorrhizae Manuscript

### 2023-01-03 DataExploratoryAnalyses.Rmd
More exploratory analyses including correlations between NEON soil microbe biodiversity and phenocam.

@@

```{r,eval=FALSE}
p <- 2 #Bud burst 
  print(NEON_phenophase_names[p])
  subDat <- read.csv(file=paste0(dataPath,'NEON_PhenologyObservations/NEON_PhenoObservationData_',gsub(" ","",NEON_phenophase_names[p]),'.csv'))
  subDat <- subDat %>% filter(phenophaseIntensity == mediumIntensity_phenophases[p])
  subDat$year <- as.numeric(subDat$year)
  subDat <- left_join(subDat,allPrecipData[,-3],by=c('siteID','date','year'))
  subDat <- left_join(subDat,allTempData,by=c('siteID','date','year'))
  subDat <- left_join(subDat,allSunlightTimes,by=c('siteID','date'))
  subDat <- subDat %>% filter(year!=2022)
  subDat$Mycorrhizal.type <- as.factor(subDat$Mycorrhizal.type)

>>>>

#For each site, do the covariates that occur with each phenophase differ between MTs? 
```{r}
p <- 2 #Bud burst 
  print(NEON_phenophase_names[p])
  subDat <- read.csv(file=paste0(dataPath,'NEON_PhenologyObservations/NEON_PhenoObservationData_',gsub(" ","",NEON_phenophase_names[p]),'.csv'))
  subDat <- subDat %>% filter(phenophaseIntensity == mediumIntensity_phenophases[p],
                              growthForm=="Deciduous broadleaf")
  subDat$year <- as.numeric(subDat$year)
  subDat <- left_join(subDat,allPrecipData[,-3],by=c('siteID','date','year'))
  subDat <- left_join(subDat,allTempData,by=c('siteID','date','year'))
  subDat <- left_join(subDat,allSunlightTimes,by=c('siteID','date'))

  ggplot(subDat,aes(as.numeric(as.character(GDD_10)),fct_reorder(siteID,decimalLatitude),fill=Mycorrhizal.type)) +
  geom_density_ridges2(scale=1)+
  theme_classic()+
  theme(axis.text.x = element_text(angle=90))+
  xlab("GDD_10 (C)")+
  ylab("Site ID (Ordered by Latitude)")+
    ggtitle("GDD_10 of Bud Burst in Deciduous Broadleaf")+
  scale_x_continuous(limits=c(0,600))

    ggplot(subDat,aes(as.numeric(as.character(GDD_5)),fct_reorder(siteID,decimalLatitude),fill=Mycorrhizal.type)) +
  geom_density_ridges2(scale=1)+
  theme_classic()+
  theme(axis.text.x = element_text(angle=90))+
  xlab("GDD_5 (C)")+
  ylab("Site ID (Ordered by Latitude)")+
    ggtitle("GDD_5 of Bud Burst in Deciduous Broadleaf")+
  scale_x_continuous(limits=c(0,600))

  ggplot(subDat,aes(as.numeric(as.character(daylength)),fct_reorder(siteID,decimalLatitude),fill=Mycorrhizal.type)) +
  geom_density_ridges2(scale=1)+
  theme_classic()+
  theme(axis.text.x = element_text(angle=90))+
  xlab("Daylength (hours)")+
  ylab("Site ID (Ordered by Latitude)")+
    ggtitle("Daylength of Bud Burst in Deciduous Broadleaf")

    ggplot(subDat,aes(as.numeric(as.character(height)),fct_reorder(siteID,decimalLatitude),fill=Mycorrhizal.type)) +
  geom_density_ridges2(scale=1)+
  theme_classic()+
  theme(axis.text.x = element_text(angle=90))+
  xlab("Plant Height")+
  ylab("Site ID (Ordered by Latitude)")+
    ggtitle("Plant Height in Deciduous Broadleaf")

What I learned: I don't really see patterns in if a MT covariates varied more for one that the other.

AM-associated plants are shorter than the EcM-associated ones so the earlier budburst is really just like because they are understory plants. Well crap.

@@

#Is phenology correlated with soil microbial biodiversity? 
```{r}
allBioDat <- read.csv(paste0(dataPath,dataName,"/soilMicrobediversityData.csv"))

allPlotIDLatLong <- matrix(nrow=0,ncol=3)
for(s in seq_along(NEON_siteNames)){
  print(NEON_siteNames[s])
  files <- dir(path=paste0(dataPath,'NEON_presence-cover-plant-CLEAN'),
               pattern=NEON_siteNames[s])
  for(f in 1:5){
    subDat <- read.csv(paste0(dataPath,'NEON_presence-cover-plant-CLEAN/',files[f]))[,c('decimalLatitude','decimalLongitude','plotID')] %>% unique()
    allPlotIDLatLong <- rbind(allPlotIDLatLong,subDat)
  }
}

allPlotIDLatLong <- allPlotIDLatLong %>% unique()
allBioDat <- left_join(allBioDat,allPlotIDLatLong,by="plotID")

phenoCamURLs <- read.csv(paste0(dataPath,'phenocamURLS.csv'))

files <- dir(path=paste0(dataPath,'phenoCamData'),pattern="dates")
allPCdat <- matrix(nrow=0,ncol=19)
for(f in seq_along(files)){
  subDat <- read.csv(paste0(dataPath,'phenoCamData/',files[f]),skip=16) %>%
    filter(gcc_value=="gcc_90")
  allPCdat <- rbind(allPCdat,subDat)

}
allPCdat_cleaned <- matrix(nrow=0,ncol=8)
subDat <- allPCdat[,c('site','veg_type','roi_id','direction',
                                                   'transition_15','transition_15_lower_ci',
                                                   'transition_15_upper_ci')]
colnames(subDat) <- c('site','veg_type','roi_id','direction',
                                                   'date','lower_ci',
                                                   'upper_ci')
subDat$transition <- 15
allPCdat_cleaned <- rbind(allPCdat_cleaned,subDat)
subDat <- allPCdat[,c('site','veg_type','roi_id','direction',
                                                   'transition_50','transition_50_lower_ci',
                                                   'transition_50_upper_ci')]
colnames(subDat) <- c('site','veg_type','roi_id','direction',
                                                   'date','lower_ci',
                                                   'upper_ci')
subDat$transition <- 50
allPCdat_cleaned <- rbind(allPCdat_cleaned,subDat)
subDat <- allPCdat[,c('site','veg_type','roi_id','direction',
                                                   'transition_85','transition_85_lower_ci',
                                                   'transition_85_upper_ci')]
colnames(subDat) <- c('site','veg_type','roi_id','direction',
                                                   'date','lower_ci',
                                                   'upper_ci')
subDat$transition <- 85
allPCdat_cleaned <- rbind(allPCdat_cleaned,subDat)
allPCdat_cleaned$siteID <- sapply(strsplit(as.character(allPCdat_cleaned$site), "[.]"), `[`, 3)
allPCdat_cleaned <- left_join(allPCdat_cleaned,allTempData,by=c('siteID','date'))
allPCdat_cleaned <- left_join(allPCdat_cleaned,allSunlightTimes,by=c('siteID','date'))
allPCdat_cleaned <- left_join(allPCdat_cleaned,allPrecipData,by=c('siteID','date','year'))

phenoCam_siteData <- read.csv(paste0(dataPath,"NEON_PhenoCam_SiteData.csv"))
phenoCam_siteData$siteID <- sapply(strsplit(as.character(phenoCam_siteData$phenoCamID), "[.]"), `[`, 3)

allPlotIDLatLong$difference <- NA
allPlotIDLatLong$siteID <- sapply(strsplit(as.character(allPlotIDLatLong$plotID), "_"), `[`, 1)
allPlotIDLatLong <- left_join(allPlotIDLatLong,phenoCam_siteData,by="siteID")
allPlotIDLatLong$direction[is.na(allPlotIDLatLong$direction)] <- "N" #Assume missing directions are North because most cameras are pointed North
allPlotIDLatLong[allPlotIDLatLong$direction=="N",'difference'] <- allPlotIDLatLong[allPlotIDLatLong$direction=="N",'decimalLatitude']-allPlotIDLatLong[allPlotIDLatLong$direction=="N",'Lat']

allPlotIDLatLong[allPlotIDLatLong$direction=="W",'difference'] <- allPlotIDLatLong[allPlotIDLatLong$direction=="W",'decimalLongitude']-allPlotIDLatLong[allPlotIDLatLong$direction=="W",'Lon']

allPlotIDLatLong[allPlotIDLatLong$direction=="NNE",'difference'] <- sqrt((allPlotIDLatLong[allPlotIDLatLong$direction=="NNE",'decimalLongitude']-allPlotIDLatLong[allPlotIDLatLong$direction=="NNE",'Lon'])**2 + (allPlotIDLatLong[allPlotIDLatLong$direction=="NNE",'decimalLatitude']-allPlotIDLatLong[allPlotIDLatLong$direction=="NNE",'Lat'])**2)

allPlotIDLatLong_filtered <- allPlotIDLatLong[allPlotIDLatLong$difference>0,] %>%
  filter(direction!="NNE") #Deal with NNE later because it's more complicated because latitude and longitude differences have to be positive

allPlotIDLatLong_filtered <- allPlotIDLatLong_filtered %>% group_by(siteID) %>%
  slice(which.min(difference))

allBioDat_plotFiltered <- allBioDat %>% 
  filter(plotID %in% allPlotIDLatLong_filtered$plotID) %>%
  mutate(year=lubridate::year(date))%>%
  group_by(year,siteID) %>%
  summarise(year_totalSpecies=sum(totalSpecies),year_totalIndividual=sum(totalIndividuals))

allBioDat_plotFiltered$biodiversity <- allBioDat_plotFiltered$year_totalSpecies/allBioDat_plotFiltered$year_totalIndividual

allPCdat_cleaned_15 <- allPCdat_cleaned %>% filter(direction=="rising",transition==15,veg_type=="DB")

allBioDat_plotFiltered_15 <- left_join(allBioDat_plotFiltered,allPCdat_cleaned_15,by=c('year','siteID'))

plot(allBioDat_plotFiltered_15$biodiversity,allBioDat_plotFiltered_15$GDD_0,pch=20)
mdl <- lm(GDD_0~biodiversity,data = allBioDat_plotFiltered_15)
abline(mdl)
summary(mdl)

What I learned: There might be some positive correlations between soil microbe biodiversity and phenocam, but there aren't enough site-years to test because soil microbe data started years before phenocam, but is 4 years delayed.

#Extra Code
```{r,eval=FALSE}
p <- 2 #Bud burst 
  print(NEON_phenophase_names[p])
  subDat <- read.csv(file=paste0(dataPath,'NEON_PhenologyObservations/NEON_PhenoObservationData_',gsub(" ","",NEON_phenophase_names[p]),'.csv'))
  subDat <- subDat %>% filter(phenophaseIntensity == mediumIntensity_phenophases[p])
  subDat$year <- as.numeric(subDat$year)
  subDat <- left_join(subDat,allPrecipData[,-3],by=c('siteID','date','year'))
  subDat <- left_join(subDat,allTempData,by=c('siteID','date','year'))
  subDat <- left_join(subDat,allSunlightTimes,by=c('siteID','date'))
  subDat <- subDat %>% filter(year!=2022)
  subDat$Mycorrhizal.type <- as.factor(subDat$Mycorrhizal.type)

### 2023-01-03 manuscriptCode_phenologyAndMycorrhizae.Rmd
More exploratory analyses including correlations between NEON soil microbe biodiversity and phenocam.

#Calculate Cummulative Weather for Each Individual phenophase intensity pair
```{r}

>>>>

#Download NEON Soil Microbe Data
```{r}
dataName="NEON_SoilMicrobeComposition"
NEON_ID='DP1.10081.001'
selectColumns <-
inFileName <- c("mcc_soilSeqVariantMetadata_16S.csv","mcc_soilSeqVariantMetadata_ITS.csv")
varName <-

downloadNEONdata(dataName=dataName,NEON_ID=NEON_ID) #Note: I had to modify this function to remove the indices for sites that didn't seem to have data. 
combineNEONdata(dataName=dataName,NEON_ID=NEON_ID,selectColumns=selectColumns,
                inFileName=inFileName,dataPath=dataPath)
```
